Instability and inequality in the COVID-19 lockdown crime drop

### 2021-03-03 draft_paper.Rmd
commit before going to NO only
osm para and crime prof gg

Individual open records categorize crime according to thirteen different notifiable offence categories (e.g. burglary, violence and sexual offences, theft from the person, vehicle crime).^[Further information on these crime categories can be obtained from https://www.police.uk/pu/contact-the-police/what-and-how-to-report/what-report/.] Records also include anti-social behaviour (ASB) which usually captures less serious offences such as nuisance behavior and is not a notifiable offence. Individual records were aggregated to create count measures by month for the thirteen notifiable offences and ASB. These are used in the initial stages of analyses to provide a background overview of nationwide trends in crime and concentration before and after lockdown. For the primary cluster analysis, we use a count measure comprised of notifiable offences excluding drug crimes. This is due to the widespread recognition that drug crime trends, particularly during the COVID-19 lockdown, largely reflect policing proactivity rather than meaningful shifts in criminal behavior [@langton2021cs]. Recognizing that aggregating data across crime types can mask variation [@andresen2017trajectories] we later decompose our findings according to the twelve remaining notifiable offences.

@@

For localized monthly analysis, notifiable offences (excluding drugs) were aggregated by Lower Super Output Area (LSOA). LSOA are a meso-level geographic units designed for the reporting of official statistics at small geographies [@onslsoa]. England and Wales is comprised of 32,844 LSOA designed to be uniform by resident population size. In 2019, LSOA housed an average of 1,700 people. Data obtained from the open police data portal (see previous section) includes a pre-assigned field stating the LSOA in which the crime occurred, as recorded by the police. Due to the spatial anonymization method used prior to data release, LSOA are the lowest level of aggregation at which we can reasonably assume spatial accuracy across multiple crime types [@tompson2015uk]. Boundaries reflect 'on the ground' characteristics such as roads and railways [@onslsoachange], and as such, we consider LSOA to represent meaningful behavioral spaces to study the opportunity structures of crime.

After removing crimes recorded by Greater Manchester Police, crimes recorded to have occurred within the Greater Manchester region, and four LSOAs which contained no crime between 2018 and 2020 (likely due to the spatial anonymization process), our final sample population for the meso-level analysis comprised 33,075 LSOAs.

@@

Analyses to examine the localized variation in the lockdown crime drop are conducted in three principal stages. First, an overview of nationwide (macro) trend is provided by crime type in terms of both absolute counts and the degree of concentration. Second, the macro-level trend is disentangled using non-parametric clustering techniques notifiable offences (excluding drugs) on the LSOA (meso) units (N = 33,075). Third, the characteristics of each cluster will be summarized in terms of their opportunity structures, spatial patterning and crime type profile. We conclude with a reflection on the findings given the crime profiles observed.

@@

Macro-level descriptives of count trends and concentration by offence type are reported for provide context from which we will subsequently unmask local (meso) variation. First, we visualize observed crime counts between February and August 2020 relative to the same periods in 2018 and 2019 by offence type. In doing so, we can observe how nationwide crime trends changed in the face of lockdown measures [see also @langton2021cs]. Second, we visualize generalized Gini coefficients [@bernasco2017more] to describe the degree of which crimes became more (or less) concentrated during lockdown at LSOA level. Again, we report figures for 2018 and 2019 as a point of reference. We use the generalized version of the Gini coefficient to account for data sparsity, as in some cases, the number of crimes fell below the number of LSOAs in the study region. Measures of counts and concentration set scene from which we can disentangle the underlying, meso-level variation.

@@

We deploy k-means using the `kml` package [@kmlpack] in R on notifiable offences (excluding drugs) counts for LSOAs in the study region (N = 33,075) between February and August 2020. To achieve a parsimonious cluster solution while minimizing the risk of missing underlying variation, we proposed potential solutions using three, four, five and six clusters respectively, choosing the final solution based on the Calinski-Criterion [@calinski1974dendrite]. For each potential solution, twenty redraws with different starting conditions were run to ensure that solutions were stable (see Appendix). This indicated that the optimal cluster solution contained six clusters.

@@

To supplement this, we visualize the spatial patterning of the cluster solutions. For brevity and simplicity, we focus on Birmingham, Liverpool, Leeds, Bradford, Sheffield and Cardiff. We have excluded Greater Manchester due to the lack of police data, and given its size, we determined Greater London to warrant an individual case study for future research. Given these exclusions, the six cities we report represent the five most populous cities in England, and the most populous city in Wales.

@@

To set the context for the localized analysis, Figure 2 visualizes crime counts between February and August 2020 for the thirteen notifiable offences, anti-social behavior, and notifiable offences excluding drugs. This descriptive analysis confirms what has been demonstrated using model-based forecasting analyses elsewhere (Langton et al., 2021). In April, the first full month of lockdown in England and Wales, we observe a widespread decline across most notifiable offences in comparison to previous years. One key exception is drug crimes, which as noted, is thought to have increased as a result of policing proactivity. Public order and violence and sexual offences experienced a dramatic decline in April, but appear to deviate little from previous years. This is due to long-term increase in these crime types (see Appendix A). The aggregate crime category, notifiable offences (excluding drugs) demonstrates a clear decline in March and April, followed by a resurgence to normality by August. This represents the 'global' trend which we will subsequently disentangle using localized analyses.

Figure 3 visualizes the overall level of concentration across these offence types before and after lockdown, and in comparison to previous years, using the generalized Gini coefficient. Here, we note that most crime types experience remarkably stable concentration, even amidst volatile and sharply changing levels in absolute counts (see Figure 2). For many crime types, the degree of concentration is almost indistinguishable compared to 2018 and 2019. Exceptions include bicycle theft, theft from the person, other theft, robbery, and (to a lesser extent) vehicle crime. These experienced a fall in concentration (i.e. crimes were less concentrated, more spread out) compared to previous years. Overall, notifiable offences (excluding drugs) experienced minimal change. There is a marginal dip in concentration during the first month of lockdown, quickly returning to normal levels of concentration as crime surged back upwards during summer.

Based on the dramatic decline and subsequent resurgence of notifiable offences (excluding drugs) during the study period, we might expect some localized variability. Measures of concentration indicate that there has been stability at the 'global' level, but the evidence-base surrounding the law of crime concentration and micro-place (in)stability would certainly give reason to suspect that this masks underlying volatility [@weisburd2015law; @andresen2017trajectories].

```{r crimeoverall, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Macro-level crime trends in England and Wales during the lockdown period in 2020."}
knitr::include_graphics("visuals/raw_counts_gg.png")
```

```{r crimegini, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Measure of concentration using the generalized Gini coefficient during the lockdown period."}
knitr::include_graphics("visuals/gini_gg.png")
```

@@

- Figure 3 visualizes the cluster solution generated by k-means, with 6-clusters identified as optimal based on the Calinski-Criterion [@calinski1974dendrite].
- Black solid line is the median count of the cluster at any given time point, with the black dotted line showing the mean.
- Violin plots have been added to convey the variation in counts at each time point. Here, we note that distributions were fairly uniform, with the mean and median representing the cluster solutions fairly well. There is little overlap between clusters.
- For reference to what trends we may have expected in a 'typical' year without the pandemic, we have added the mean and median trend lines for each cluster during 2018 (in blue) and 2019 (in red).
- This demonstrates that the clusters identified using the 2020 data were distinct and meaningful even in previous years, and highlights the gap between previous years and what has been observed during the pandemic.
- Overall, we note that most LSOA were remarkably stable during the pandemic. Clusters A and B can be described and 'low crime and stable', exhibiting fairly low counts throughout the study period and across years. Together, these clusters comprise 89% of LSOAs in England and Wales.
- That said, there is some deviation from previous years in April, in alignment with expectations. There is a slight dip in the mean and median crime count, particularly in Cluster B.
- The third largest cluster, cluster C, comprises 10% of all LSOA in the country. This cluster experienced a more prominent dip in crime, along with cluster D (1.6% of LSOAs). Together, we might describe these LSOAs as 'mid-crime, mid-drop'. 
- However, the 'action' appears to be occurring amongst a small subset of LSOAs, contained within clusters E and F. Together, these clusters comprise only 0.33% of all LSOA in the country (N = 110), however they are 'high crime, major drop'.
- Interestingly, clusters E and F represent the only clusters which did not return back to normality by August (if we assume that 2018 and 2019 represent a trend which we might have otherwise expected).
- Overall, we note that despite the dramatic nationwide trend in notifiable offences (excluding drugs), there has been remarkable stability in local areas, with most change being driven by a small number of areas.

```{r kmeansnotoff, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "K-means cluster solutions for LSOA notifiable offences (excluding drugs). Distributions refer to 2020 only.", out.width= "90%"}
knitr::include_graphics("visuals/kmeans_violin_full_gg.png")
```

@@

- To further unpick the contribution of these clusters to the nationwide lockdown crime drop, Figure 4 plots the monthly change in counts of each cluster.
- Alongside these raw counts, we also include the percentage of total absolute change (i.e. positive or negative) attributable to each cluster.
- We use this to identify which clusters drove the initial decline and subsequent resurgence nationwide.
- As expected, the vast majority of change occurred between March and April.
- However, Figure 4 demonstrates that this substantial decline did not occur equitably across the country. Consistent with the cluster solutions trends in Figure 3, it is clear that a small number of areas drove the trend.
- For instance, clusters A and B, which comprise 89% of LSOAs in the country, accounted for only 47% of the decline between March and April.
- By contrast, clusters E and F, which together comprised only 0.33% of LSOA (N = 110) contributed to 15% of the nationwide drop between months. This a disproportionately large impact of the macro-level trend, and further supports the observation that most areas remained fairly stable.
- Interestingly, this picture shifts upon the beginning of the resurgence between April and May.
- Cluster A continues to contribute a disproportionately small amount to nationwide change, but cluster B contributes to 42% of the increase, despite only comprising 31% of LSOA.
- The contribution of the smallest clusters E and F is remarkably small compared to the initial decline, 2% and 0% respectively. This can be largely attributed to the flattening of their crime count trends between these months. In other words, following their initially steep decline between March and April, crime remained reasonably unchanged between April and May in these areas.
- Overall, this pattern continues throughout the spring convergence back to normality, with some notable changes. 
- Following the period of stabilization, clusters E and F gather pace during the resurgence, as does cluster D. For instance, between June and July, these three clusters account for nearly a third of the nationwide increase, despite comprising less than 2% of total LSOA.
- July to August represents an interesting period. Until that point, each cluster, despite their vastly differing relative contributions to change, were directionally homogeneous. In other words, every cluster declined between February and April, and then increased between April and July.
- Between July and August, clusters began to diverge. Nationwide, there was a marginal increase in notifiable offence counts (see Figure 1), but this masks a great deal of localized variability.
- The two large, low crime clusters (A and B) actually declined again between July and August, while the higher crime clusters continued to increase. This is suggestive of a local change in opportunities as lockdown rules were relaxed - change which has been aggregated away at the macro-level.
- In other words, low crime areas mitigated against further nationwide increases which may, if they had continued, actually resulted in a higher crime count in August than observed in 2019.

```{r changeline, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Counts and percentage of nationwide change between months attributable to each cluster.", out.width= "100%"}
knitr::include_graphics("visuals/change_k6_gg.png")
```

## Spatial distribution of clusters

- Link to interactive map

```{r  maps, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Spatial distribution of clusters by major urban conurbation.", out.width= "100%"}
knitr::include_graphics("visuals/maps_gg.png")
```

@@

- We have now identified that, amidst dramatic nationwide change in notifiable offences during lockdown, and stability in overall crime concentrations, there has been considerable underlying volatility, with a small number of local areas driving the macro-level trends.
- From an opportunity perspective, we expect the clusters to have differing opportunity structures.
- For instance, LSOA which were typically 'high crime' in 2018 and 2019, and then experienced a dramatic fall in crime upon lockdown commencement, are likely to have a large volume of criminal opportunities which were closed off following restrictions on mobility.
- On the other hand, areas which were typically low crime, and remained low crime, are likely to have a vastly different opportunity structure: places with minimal attractors or generators to begin with are unlikely to have changed dramatically during lockdown.
- Table 1 demonstrates this point using the data obtained from Open Street Map.
- Clusters A and B, the 'low crime and stable' LSOAs, tend to have few major opportunities for crime. Few pubs, nightclubs or restaurants to generate a violence and theft-prone nightlife, minimal shopping opportunities (and thus less chance of shoplifting), limited public transport (less theft) and less bicycle parking (less bicycle theft).
- By contrast, the 'high crime major drop' clusters (e.g. E and F) are major nightlife areas, have considerable shopping opportunities, often have expansive public transport networks, and have many bicycle parking facilities.
- These findings are consistent with what is observed using the interactive map made available as part of this paper (link here).

@@

### Crime type profile

@@

@@

```{r longterm, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Select crime types with long-term increases in counts.", out.width= "90%"}
knitr::include_graphics("visuals/long_term_gg.png")
```

```{r crimeprofile, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Crime type characteristics of each cluster solution.", out.width= "90%"}
knitr::include_graphics("visuals/traj_crimes_monthly_k6_gg.png")
```

>>>>

Individual open records categorize crime according to thirteen different notifiable offence categories (e.g. burglary, violence and sexual offences, theft from the person, vehicle crime).^[Further information on these crime categories can be obtained from https://www.police.uk/pu/contact-the-police/what-and-how-to-report/what-report/.] Records also include anti-social behaviour (ASB) which usually captures less serious offences such as nuisance behavior and is not a notifiable offence. Individual records were aggregated to create a count measure for notifiable offences (excluding drugs) by month either at the nationwide (macro) level, or the localized (meso) level, as detailed in the next section. There widespread recognition that drug crime trends, particularly during the COVID-19 lockdown, largely reflect policing proactivity rather than meaningful shifts in criminal behavior [@langton2021cs]. Recognizing that aggregating data across crime types can mask variation [@andresen2017trajectories] we later decompose our main findings according to the twelve remaining notifiable offences, and refer to additional analyses in the Appendix.

@@

For nationwide (macro) descriptive statistics, which set the background for the main analyses, we use the geographic region of England and Wales, noting the exclusion of Greater Manchester. For localized (meso) analysis, we aggregate offences to Lower Super Output Area (LSOA). LSOA are a meso-level geographic units designed for the reporting of official statistics at small geographies [@onslsoa]. England and Wales is comprised of 32,844 LSOA designed to be uniform by resident population size. In 2019, LSOA housed an average of 1,700 people. Data obtained from the open police data portal (see previous section) includes a pre-assigned field stating the LSOA in which the crime occurred, as recorded by the police. Due to the spatial anonymization method used prior to data release, LSOA are the lowest level of aggregation at which we can reasonably assume spatial accuracy across multiple crime types [@tompson2015uk]. Boundaries reflect 'on the ground' characteristics such as roads and railways [@onslsoachange], and as such, we consider LSOA to represent meaningful behavioral spaces to study the opportunity structures of crime. After removing crimes recorded by Greater Manchester Police, crimes recorded to have occurred within the Greater Manchester region, and four LSOAs which contained no crime between 2018 and 2020 (likely due to the spatial anonymization process), our final sample population for the meso-level analysis comprised 33,075 LSOA.

@@

Analyses to examine the localized variation in the lockdown crime drop are conducted in three principal stages. First, an overview of nationwide (macro) trend is provided in terms of both absolute counts and the degree of concentration. Second, the macro-level trend is disentangled using non-parametric clustering techniques on the LSOA (meso) units (N = 33,075). Third, the characteristics of each cluster will be summarized in terms of their opportunity structures, spatial patterning and crime type profile.

@@

Macro-level descriptives of count trends and concentration for notifiable offences (excluding drugs) are reported for provide context from which we will subsequently unmask local (meso) variation. First, we visualize observed counts between February and August 2020 relative to the same periods in 2018 and 2019. In doing so, we can observe how nationwide crime trends changed in the face of lockdown measures [see also @langton2021cs]. Second, we visualize generalized Gini coefficients [@bernasco2017more] to describe the degree of which offences became more (or less) concentrated during lockdown at LSOA level. Again, we report figures for 2018 and 2019 as a point of reference. We use the generalized version of the Gini coefficient to account for data sparsity, as for some crime types reported in the Appendix, the number of crimes falls below the number of LSOAs in the study region. Measures of counts and concentration set scene from which we can disentangle the underlying, meso-level variation.

@@

We deploy k-means using the `kml` package [@kmlpack] in R on notifiable offences (excluding drugs) counts for LSOA in the study region (N = 33,075) between February and August 2020. To achieve a parsimonious cluster solution while minimizing the risk of missing underlying variation, we proposed potential solutions using three, four, five and six clusters respectively, choosing the final solution based on the Calinski-Criterion [@calinski1974dendrite]. For each potential solution, twenty redraws with different starting conditions were run to ensure that solutions were stable (see Appendix).

@@

To supplement this, we visualize the spatial patterning of the cluster solutions. For brevity and simplicity, we focus on Birmingham, Liverpool, Leeds, Bradford, Sheffield and Cardiff. We have excluded Greater Manchester due to the lack of police data, and given its size, we determined Greater London to warrant an individual case study for future research. Given these exclusions, the six cities we report represent the five most populous cities in England, and the most populous city in Wales. Study regions are defined by based on the city names appearing in LSOA name.

@@

To set the context for the localized analysis, Figure 2 visualizes crime counts between February and August 2020 for notifiable offences excluding drugs before and after lockdown. This descriptive analysis broadly confirms what has been demonstrated using model-based forecasting analyses elsewhere (Langton et al., 2021) although we note that this descriptive comparison does not account for long-term trends. In April, the first full month of lockdown in England and Wales, we observe a nationwide decline in notifiable offences in comparison to previous years. Upon the relaxation of lockdown rules, crime resurged back to normality, and by August, crime had returned to within a range we might have expected without the nationwide lockdown. This trend represents the 'global' trend which we will subsequently disentangle using localized analyses.

```{r nocounts, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Notitifable offence (excluding drugs) counts in England and Wales during the lockdown period.", fig.align='center'}
knitr::include_graphics("visuals/no_raw_counts_gg.png")
```

Figure 3 visualizes the overall level of concentration before and after lockdown, and in comparison to previous years, using the generalized Gini coefficient. Here, it is clear that notifiable offences, even amidst volatile and sharply changing levels in absolute counts (see Figure 2), experienced remarkably stable levels of concentration. There is a marginal dip in concentration during April, the first month of lockdown. This suggests that as crimes fell, those were occurring were _less_ concentrated. That is, crimes occurring during the height of lockdown were more spread out that we might have otherwise expected. That said, concentration quickly returned to normal levels.

```{r nogini, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Monthly generalized Gini cofficient trends for notitifable offences (excluding drugs) in England and Wales during the lockdown period.", fig.align='center'}
knitr::include_graphics("visuals/no_gini_gg.png")
```

Based on the dramatic decline and subsequent resurgence of notifiable offences during the study period, we might expect some localized variability. Measures of concentration indicate that there has been stability at the 'global' level, but the evidence-base surrounding the law of crime concentration and micro-place (in)stability would certainly give reason to suspect that this masks underlying volatility [@weisburd2015law; @andresen2017trajectories]. With this in mind, we now report the localized, meso-level analysis which unpicks the macro-level trend.

@@

We deploy non-parametric k-means clustering on LSOA (meso) level geographic units (N = 33,075) comprising England and Wales to decompose the macro-level 'decline and resurgence' observed between February and August, 2020. Figure 3 visualizes the resulting 6-cluster solution determined by the optimal Calinski-Criterion statistic [@calinski1974dendrite]. Solid black lines represent the median count for each cluster at any given time point, with the black dotted line showing the mean. Violin plots have been added to convey variation around these points at each time point. These suggest that the mean and median point statistics summarize the underlying data reasonably well, and indicate that clusters are distinct from one another, with little overlap. Additional lines have been added to convey each clusters' mean and median trend in 2018 and 2019 respectively. These trends suggest that the clusters identified using the 2020 study period were distinct and meaningful even in previous years, and provide a relative baseline from which we can compare lockdown trends.

Overall, we note that most LSOA were remarkably stable during the pandemic. Clusters A and B could be described and 'low crime and stable', exhibiting fairly low counts throughout the study period and across years. Together, these clusters comprise 89% of LSOA in England and Wales. Even amidst the stark macro-level decline in notifiable offences (see Figure 2) LSOA in these clusters only experienced marginal average dips in crime. The third largest cluster, cluster C, comprises 10% of LSOA in England and Wales. LSOA in this cluster experienced a more prominent dip in crime, along with cluster D (1.6% of LSOA). Together, we might describe these LSOAs as 'mid-crime, mid-drop'. In both cases, there is a clear deviation from previous years. Notifiable offences fall between March and April, and then began to converge back to levels observed in previous years.

```{r kmeansnotoff, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "K-means cluster solutions for LSOA notifiable offences (excluding drugs). Distributions refer to 2020 only.", out.width= "90%", fig.align='center'}
knitr::include_graphics("visuals/kmeans_violin_full_gg.png")
```

That said, most 'action' appears to be occurring amongst a small subset of LSOA. Clusters E and F collectively comprise only 0.33% (N = 110) of LSOA in England and Wales. Yet, their crime counts are much higher, and the decline between March and April is considerable. We might therefore describe these clusters as 'high crime, major drop': LSOA with plentiful opportunities for crime in typical times, and in turn, LSOA which are most sensitive to the restriction in opportunities which followed after the imposition of lockdown. Interestingly, clusters E and F represent the only clusters which remained below pre-pandemic levels by August. In both cases, the average counts in February 2020 were _higher_ than those observed in previous years, but following the sharp decline and subsequent increase by August, these averages were _lower_ than 2018 and 2019.

@@

To further unpick the contribution of these clusters to the nationwide lockdown crime drop, Figure 4 plots the month-by-month change in counts. Alongside these raw counts, we also include the percentage of total absolute change (i.e. positive or negative) attributable to each cluster. We can use this visual to identify which clusters drove the initial decline and subsequent resurgence nationwide.

As expected, the vast majority of change across all clusters occurred between March and April. Between these months, notifiable offences experienced a dramatic fall nationwide. That said, the figure demonstrates that this decline did not occur equitably across local areas. Consistent with the cluster solutions trends in Figure 3, a small number of LSOA (meso) units appear to have contributed disproportionately to the nationwide (macro) trend. For instance, clusters A and B, which comprise 58% and 31% of LSOA in the country, accounted for only 20% and 27% of the total decline between March and April. By contrast, clusters E and F, which together comprised only 0.33% of LSOA (N = 110) contributed to 15% of the nationwide drop in these months. Again, this suggests that most areas remarkably stable, given the nationwide volatility, with just a handful of areas driving the decline the nationwide lockdown crime drop.

This picture shifts slightly upon the beginning of the resurgence between April and May. Clusters A and B continue to contribute a disproportionately _small_ amount to nationwide change, but the proportion increases, with each contributing 24% and 42% to the increase respectively. The contribution of the smallest clusters E and F is small compared to the initial decline, 2% and 0% (when rounded), respectively. This can be largely attributed to the flattening of their crime count trends between these months. In other words, following their initially steep decline between March and April, crime remained reasonably unchanged between April and May in the 'high crime, major drop' areas.

Following the period of stabilization, clusters E and F gather pace during the resurgence, as does cluster D. For instance, between June and July, these three clusters collectively account for 27% of the nationwide increase, despite comprising less than 2% of LSOA in the country. July to August represents an interesting period. Until that point, each cluster, despite their vastly differing relative contributions to change, were directionally homogeneous. In other words, every cluster declined between February and April, and then increased between April and July. Between July and August, clusters began to diverge. Nationwide, there was a marginal increase in notifiable offence counts (see Figure 1), but this clearly masks a great deal of localized variability.

The 'low crime and stable' clusters (A and B) actually _declined again_ between July and August, and the higher crime clusters continued to increase. The change is almost imperceptible when viewing average counts in these LSOA (see Figure 3) but the sheer size of these units, collectively comprising 89% of England and Wales, has a major impact on nationwide trends. This is suggestive of highly localized change in opportunities as lockdown rules were relaxed: change which is aggregated away at the macro-level. Stable, low crime areas mitigated against further nationwide increases which may, if they had continued, actually resulted in a higher crime count in August than observed in previous years.

```{r changeline, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Counts and percentage of nationwide change between months attributable to each cluster.", out.width= "100%", fig.align='center'}
knitr::include_graphics("visuals/change_k6_gg.png")
```

@@

# Spatial distribution

The provide local context to the cluster solutions identified through the k-means clustering, we visualize the spatial patterning of groupings for six major cities in England and Wales (see Figure 5). We find distinct geographic pattern to the clusters identified. Here, it is worth noting that the Bradford region contains the city of Bradford, but also the satellite town of Keighley to the north west. The cluster map thus reflects these two urban centers. Without exception, the city centres and commercial districts of each city are characterised by 'high crime major drop' clusters E and F, which in turn often neighbor the 'mid crime mid drop' cluster D. By contrast, the 'low crime and stable' clusters tend to sit outside of the city centres: cluster B in the suburbs and the sparsely populated cluster A LSOA on the periphery of urban conurbations. For a detailed investigation of the spatial clusters identified for the whole of England and Wales, we refer readers to an openly available online map (anonymized link here).

```{r  maps, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Spatial distribution of clusters by major urban conurbation.", out.width= "100%", fig.align='center'}
knitr::include_graphics("visuals/maps_gg.png")
```

# Opportunity structure

We have now identified that, amidst dramatic nationwide change in notifiable offences during lockdown, and stability in overall crime concentrations, there has been considerable underlying volatility, with a small number of local areas driving the macro-level trend. From a theoretical perspective, we expect the clusters to have differing opportunity structures. Table 1 reports a series of descriptive statistics based on the opportunity structure of each cluster using the nationwide data obtained from the Open Street Map API. These findings largely support our expectations, namely, that large, stable clusters have few opportunities for crime, while high crime areas which were responsible for a disproportionately large amount of the crime drop have plentiful features which facilitate crime.

By way of example, there are sparse opportunities for crime in the largest and most stable cluster A. Most LSOA in this cluster have no nightlife facilities, no shops and no bicycle parking. Public transport is available but it is not common: the median LSOA in cluster A only contained four bus stops or railways stations. As we move along to medium and high crime clusters, which had higher pre-existing crime levels, and thus steeper declines during lockdown, these counts markedly increase. For instance, LSOA in cluster E contain a median of 21 nightlife facilities, 28 shops, 24 public transport nodes and 13 bicycle parking spaces. These counts increase further for cluster F, although we note that the cluster contains only nine LSOA.

@@

# Crime type profile

@@

```{r crimeprofile, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Crime type characteristics of each cluster solution.", out.width= "100%", fig.align='center'}
knitr::include_graphics("visuals/traj_crimes_monthly_k6_gg.png")
```

@@

```{r crimeoverall, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Counts by offence type in England and Wales during the lockdown period.", fig.align='center'}
knitr::include_graphics("visuals/raw_counts_gg.png")
```

```{r crimegini, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Monthly generalized Gini cofficient trends by offence type in England and Wales during the lockdown period.", fig.align='center'}
knitr::include_graphics("visuals/gini_gg.png")
```

```{r longterm, fig.pos = 'H', message=F, comment=F, warning=F, fig.cap = "Select crime types with long-term increases in counts.", out.width= "90%", fig.align='center'}
knitr::include_graphics("visuals/long_term_gg.png")
```
